### 2012-11-10 paper/isca40format/InvariantsHopeful.tex
will be removing evict req

\newcommand{\Req}[3]{\threeAngle{Req}{#1}{#2}{#3}}
\newcommand{\Resp}[3]{\threeAngle{Resp}{#1}{#2}{#3}}
\newcommand{\Inv}{\text{$Inv$\;}}

The values of the coherence state (and hence the directory state) belong to a
totally ordered set, \ie a set in which a \emph{less than} ($<$) relation is
defined between any two set elements and this relation is transitive. If an
address is not present, then the coherence state corresponding to that address
is the lowest value in the set (which we denote by \Inv. For instance, in an
MSI protocol, the values of the coherence state are from the set $\{M, S, I\}$
where $I < S$, $S < M$ and $I < M$; if an address is not present, the state
corresponding to that address is $I$. Our abstract protocol does not assign any
semantics or meaning to the values of the coherence states. In a concrete cache
coherence protocol, the values of the states are associated with permissions
for reading and writing, and the state with fewer permissions is $<$ the state
with more permissions.

@@

\begin{figure}\centering
\begin{tabularx}{\linewidth}{|l|X|}
\hline
$isPresent(a)$ & Is line for address $a$ present in this node?\\
\hline
$isWait(n, a)$ & Is there a pending response from node $n$ (either parent $P$
or a child) for address $a$?\\
\hline
$isReplacing(a)$ & Is address $a$ currently being evicted?\\
\hline
$isEnterPending(a)$ & Is there any (child) request for address $a$ waiting to get
a cache line entry in the cache?\\
\hline
$isHandling(n, a)$ & Is there any request for address $a$ from node $n$ (either
parent $P$ or a child) that has started to be handled?\\
\hline
%$replacingAddr(a)$ & The address that is going to replace the line
%corresponding to address $a$\\
%\hline
%$cWaitState(c, a)$ & The state to which the node has requested its child $c$ to
%downgrade to and hence waiting for a response\\
%\hline
\end{tabularx}
\caption{Information that can be computed for each address (typically from the
MSHR)}
\label{table:functions}
\end{figure}

\begin{figure}
\begin{tabularx}{\linewidth}{|l|l|X|}
\hline
\multicolumn{3}{|c|}{\Req{n}{a}{x}}\\
\hline
\multirow{4}{*}{$n = P$} & Incoming & Request from parent for changing my
                                      $state[a]$ to $x$\\
\cline{2-3}
                         & Outgoing & Request to parent for permission to change
                                      my $state[a]$ to $x$\\
\hline
\multirow{4}{*}{$n \neq P$} & Incoming & Request from child $n$ for permission
                                         to change $n$'s $state[a]$ to $x$\\
\cline{2-3}
                            & Outgoing & Request to child $n$ to change $n$'s
                                         $state[a]$ to $x$\\
\hline
\multicolumn{3}{|c|}{\Resp{n}{a}{x}}\\
\hline
\multirow{4}{*}{$n = P$} & Incoming & Response from parent granting permission
                                      to change my $state[a]$ to $x$\\
\cline{2-3}
                         & Outgoing & Response to parent indicating that my
                                      $state[a]$ has changed to $x$\\
\hline
\multirow{4}{*}{$n \neq P$} & Incoming & Response from child indicating that
                                         $n$'s $state[a]$ has changed to $x$\\
\cline{2-3}
                            & Outgoing & Response to child $n$ granting permission
                                         to change $n$'s $state[a]$ to $x$\\
\hline
\end{tabularx}
\caption{Incoming and outgoing messages}
\label{table:messages}
\end{figure}

\begin{figure}
\begin{tabularx}{\linewidth}{|X|}
\hline
A response for address $a$ can be sent only if the line for $a$ is present in
the current node\\
\hline
\multicolumn{1}{|c|}{Response to child $c$ for address $a$}\\
\hline
\begin{enumerate}
\item A response to child $c$ for address $a$ can be sent only if a
request from child $c$ for $a$ is being handled
\item Response \Resp{c}{a}{x} can be sent only if $x > dir[a][c]$, $\forall c',
c' \neq c \rightarrow isCompatible(x, dir[a][c'])$ and  $x \le state[a]$
\end{enumerate}\\
\hline
\multicolumn{1}{|c|}{Response to parent $P$ for address $a$}\\
\hline
\begin{enumerate}
\item If there is a pending response from the parent for address $a$, a
response to the parent for $a$ can be sent only if a request from
the parent for $a$ is being handled
\item If there is no pending response from the parent for address $a$, a
response to the parent for $a$ can be sent any time
\item Response \Resp{P}{a}{x} can be sent only if $x < state[a]$ and $\forall
c, x \ge dir[a][c]$
\end{enumerate}\\
\hline
\end{tabularx}
\caption{Sending a response: Preconditions}
\label{sendRespPre}
\end{figure}

\begin{figure}
\begin{tabularx}{\linewidth}{|X|}
\hline
A response for address $a$ can be sent only if the line for $a$ is present in
the current node\\
\hline
\multicolumn{1}{|c|}{Response to child $c$ for address $a$}\\
\hline
\begin{enumerate}
\item[] On sending \Resp{c}{a}{x}, $dir[a][c]$ changes to $x$
\end{enumerate}\\
\hline
\multicolumn{1}{|c|}{Response to parent $P$ for address $a$}\\
\hline
\begin{enumerate}
\item[] On sending \Resp{P}{a}{x}, $state[a]$ changes to $x$
\end{enumerate}\\
\hline
\end{tabularx}
\caption{Sending a response: Effects}
\label{sendRespEff}
\end{figure}

\begin{figure}
\begin{tabularx}{\linewidth}{|X|}
\hline
\multicolumn{1}{|c|}{On receiving a response from child $c$ for address $a$}\\
\hline
\begin{enumerate}
\item $dir[a][c]$ changes to $x$
\item Response \Resp{c}{a}{x} is removed
\end{enumerate}\\
\hline
\multicolumn{1}{|c|}{On receiving a response from parent $c$ for address $a$}\\
\hline
\begin{enumerate}
\item $state[a]$ changes to $x$
\item Response \Resp{P}{a}{x} is removed
\end{enumerate}\\
\hline
\end{tabularx}
\caption{Receiving a response: Effects}
\label{recvResp}
\end{figure}

@@

\begin{figure}
\begin{tabularx}{\linewidth}{|X|}
\hline
\multicolumn{1}{|c|}{Request to child $c$ for address $a$}\\
\hline
\begin{enumerate}
\item A request can be sent to child $c$ for address $a$ only if there is no
pending response from child $c$ for $a$
\item Request \Req{c}{a}{x} can be sent only if $x < dir[a][c]$
\end{enumerate}\\
\hline
\multicolumn{1}{|c|}{Request to parent $P$ for address $a$}\\
\hline
\begin{enumerate}
\item A request can be sent to the parent for address $a$ only if
  \begin{enumerate}
  \item there is no pending response from the parent for address $a$, and
  \item line in address $a$ is not being replaced
  \end{enumerate}
\item Request \Req{P}{a}{x} can be sent only if $x > state[a]$
\end{enumerate}\\
\hline
\end{tabularx}
\caption{Sending a request: Preconditions}
\label{sendReq}
\end{figure}

\begin{figure}
\begin{tabularx}{\linewidth}{|X|}
\hline
\multicolumn{1}{|c|}{Request \Req{c}{a}{x} from child $c$ is being handled}\\
\hline
\begin{itemize}
\item[] If $x \le state[a] \wedge \forall c', c' \neq c \rightarrow isCompatible(x, dir[a][c])$\newline
 then the node must send a response \Resp{c}{a}{x}
to $c$
\end{itemize}\\
\hline
\multicolumn{1}{|c|}{Address $a$ is marked as being replaced}\\
\hline
\begin{enumerate}
\item[] If $\forall c, dir[a][c] \le \Inv$, then the node must send a response
\Resp{P}{a}{\Inv} to the parent
\end{enumerate}\\
\hline
\multicolumn{1}{|c|}{Request \Req{P}{a}{x} from parent $P$ is being
handled}\\
\hline
\begin{enumerate}
\item[] If $\forall c, dir[a][c] \le x$, then the node must send a response
\Resp{P}{a}{x} to the parent
\end{enumerate}\\
\hline
\end{tabularx}
\caption{Forced responses: Effects}
\label{respForReq}
\end{figure}

\begin{figure}
\begin{tabularx}{\linewidth}{|X|}
\hline
\multicolumn{1}{|c|}{Request \Req{c}{a}{x} from child $c$ is received}\\
\hline
\begin{enumerate}
\item If $a$ is present in the node, handling of request \Req{c}{a}{x} can be
started only if
  \begin{enumerate}
  \item no request \Req{n}{a}{x} from any node $n$ for address $a$ is being
  handled
  \item address $a$ is not currently being replaced
  \end{enumerate}
\item If $a$ is not present in the node, handling of request \Req{c}{a}{x} can be
started only if there is a line with address $a'$ such that
  \begin{enumerate}
  \item no request \Req{n}{a'}{x} from any node $n$ for address $a'$ is being
  handled
  \item address $a'$ is not currently being evicted
  \end{enumerate}
\Req{n}{a'}{x} has started
\end{enumerate}\\
\hline
\multicolumn{1}{|c|}{Request \Req{P}{a}{x} from parent $P$ is received}\\
\hline
\begin{enumerate}
\item[] Handling of request \Req{P}{a}{x} can be started only if there are no
pending responses from any child for address $a$
\end{enumerate}\\
\hline
\end{tabularx}
\caption{Start handling a request: Preconditions}
\label{respForReqPre}
\end{figure}

\begin{figure}
\begin{tabularx}{\linewidth}{|X|}
\hline
\multicolumn{1}{|c|}{Request \Req{c}{a}{x} from child $c$ is received}\\
\hline
\begin{enumerate}
\item Start handling the request \Req{c}{a}{x}
\item If $a$ is not present in the node, get a line with address $a'$ such that
  \begin{enumerate}
  \item no request \Req{n}{a'}{x} from any node $n$ for address $a'$ is being
  handled
  \item address $a'$ is not currently being evicted
  \end{enumerate}
Mark $a'$ as being evicted
\end{enumerate}\\
\hline
\multicolumn{1}{|c|}{Request \Req{P}{a}{x} from parent $P$ is received}\\
\hline
\begin{enumerate}
\item[] Start handling the request \Req{P}{a}{x}
\end{enumerate}\\
\hline
\end{tabularx}
\caption{Start handling a request: Effects}
\label{respForReqEff}
\end{figure}

\begin{figure}
\begin{tabularx}{\linewidth}{|X|}
\hline
\multicolumn{1}{|c|}{Request \Req{c}{a}{x} from child $c$ is being handled}\\
\hline
\begin{enumerate}
\item[] If $x \le dir[a][c]$, remove the request
\end{enumerate}\\
\hline
\multicolumn{1}{|c|}{Request \Req{P}{a}{x} from parent $P$ is being
handled}\\
\hline
\begin{enumerate}
\item[] If $x \ge state[a]$, remove the request
\end{enumerate}\\
\hline
\end{tabularx}
\caption{Finish handling a request: Effects}
\label{finishHandle}
\end{figure}

\begin{figure}
\begin{tabularx}{\linewidth}{|X|}
\hline
Request \Req{c}{a}{x} from child $c$ is being handled, and address $a'$ is
replaced by $a$\\
\hline
\begin{enumerate}
\item If $state[a'] <= \Inv$, replace address of the line for $a'$ as $a$,
$state[a']$ as \Inv and $\forall c, dir(c, a)$ as \Inv
\item Unmark $a'$ as being evicted
\end{enumerate}\\
\hline
\end{tabularx}
\caption{Finish replacing line: Effects}
\end{figure}

\begin{figure}
\begin{tabularx}{\linewidth}{|X|}
\hline
\begin{enumerate}
\item[] If the state or dir of an address (either the requested address, or the
evicted address in case of a replacement) for some node $n$ is not satisfactory,
and there are no pending responses from $n$, send a request to make the state
or dir satisfactory
\end{enumerate}\\
\hline
\end{tabularx}
\caption{Handling of a request: Effects}
\label{handling}
\end{figure}

\begin{figure}
\begin{tabularx}{\linewidth}{|X|}
\hline
\begin{enumerate}
\item[] The effects during sending a response or sending a request is atomic,
\ie if there is no space available in the sending buffer, then the whole action
is retried later
\end{enumerate}\\
\hline
\end{tabularx}
\caption{Order of receiving messages}
\label{atomicity}
\end{figure}

>>>>

\newcommand{\Req}[3]{\threeAngle{Req}{#1}{#2}{#3}}
\newcommand{\Resp}[3]{\threeAngle{Resp}{#1}{#2}{#3}}
\newcommand{\Inv}{\text{$Inv$}{}}
\newcommand{\Dirty}{\text{$Dirty$}{}}
\newcommand{\True}{\text{$True$}{}}
\newcommand{\False}{\text{$False$}{}}

The values of the coherence state (and hence the directory state) belong to a
totally ordered set, \ie a set in which a \emph{less than} ($<$) relation is
defined between any two set elements and this relation is transitive. If an
address is not present, then the coherence state corresponding to that address
is the lowest value in the set (which we denote by \Inv). For instance, in an
MSI protocol, the values of the coherence state are from the set $\{M, S, I\}$
where $I < S$, $S < M$ and $I < M$; if an address is not present, the state
corresponding to that address is $I$. Our abstract protocol does not assign any
semantics or meaning to the values of the coherence states. In a concrete cache
coherence protocol, the values of the states are associated with permissions
for reading and writing, and the state with fewer permissions is less than the
state with more permissions.

@@

\begin{figure}\centering
\begin{tabularx}{\linewidth}{|l|X|}
\hline
$isPresent(a)$ & Is line for address $a$ present in this node?\\
\hline
$isWait(n, a)$ & Is there a pending response from node $n$ (either parent $P$
or a child) for address $a$?\\
\hline
$isEvicting(a)$ & Is address $a$ currently being evicted?\\
\hline
%$isEnterPending(a)$ & Is there any (child) request for address $a$ waiting to get
%a cache line entry in the cache?\\
%\hline
$isHandling(n, a)$ & Is there any request for address $a$ from node $n$ (either
parent $P$ or a child) that is being handled?\\
\hline
%$replacingAddr(a)$ & The address that is going to replace the line
%corresponding to address $a$\\
%\hline
%$cWaitState(c, a)$ & The state to which the node has requested its child $c$ to
%downgrade to and hence waiting for a response\\
%\hline
\end{tabularx}
\caption{Information that can be computed for each address (typically from the
MSHR)}
\label{table:functions}
\end{figure}

\begin{figure}
\begin{tabularx}{\linewidth}{|cX|}
\hline
\multicolumn{2}{|c|}{\Req{n}{a}{x}, \Resp{n}{a}{x}}\\
\hline
$n:$ & Source or destination node ID (parent $P$, or a child)\\
$a:$ & Address of the line corresponding to the request\\
$x:$ & Value from the set of coherence states\\
\hline
\end{tabularx}
\caption{Incoming and outgoing message}
\label{message}
\end{figure}

%\begin{figure}
%\begin{tabularx}{\linewidth}{|l|l|X|}
%\hline
%\multicolumn{3}{|c|}{\Req{n}{a}{x}}\\
%\hline
%\multirow{4}{*}{$n = P$} & Incoming & Request from parent for changing my
%                                      $state[a]$ to $x$\\
%\cline{2-3}
%                         & Outgoing & Request to parent for permission to change
%                                      my $state[a]$ to $x$\\
%\hline
%\multirow{4}{*}{$n \neq P$} & Incoming & Request from child $n$ for permission
%                                         to change $n$'s $state[a]$ to $x$\\
%\cline{2-3}
%                            & Outgoing & Request to child $n$ to change $n$'s
%                                         $state[a]$ to $x$\\
%\hline
%\multicolumn{3}{|c|}{\Resp{n}{a}{x}}\\
%\hline
%\multirow{4}{*}{$n = P$} & Incoming & Response from parent granting permission
%                                      to change my $state[a]$ to $x$\\
%\cline{2-3}
%                         & Outgoing & Response to parent indicating that my
%                                      $state[a]$ has changed to $x$\\
%\hline
%\multirow{4}{*}{$n \neq P$} & Incoming & Response from child indicating that
%                                         $n$'s $state[a]$ has changed to $x$\\
%\cline{2-3}
%                            & Outgoing & Response to child $n$ granting permission
%                                         to change $n$'s $state[a]$ to $x$\\
%\hline
%\end{tabularx}
%\caption{Incoming and outgoing messages}
%\label{table:messages}
%\end{figure}

\begin{figure}
\begin{tabularx}{\linewidth}{|X|}
\hline
\begin{enumerate}
\item[] A response for address $a$ can be sent only if the line for $a$ is present in
the current node, \ie $isPresent(a)$
\end{enumerate}\\
\hline
\multicolumn{1}{|c|}{Response to child $c$ for address $a$}\\
\hline
\begin{enumerate}
\item A response to child $c$ for address $a$ can be sent only if a
request from child $c$ for $a$ is being handled, \ie $isHandling(c, a)$
\item Response \Resp{c}{a}{x} can be sent only if $x > dir[a][c]$
% $\forall c', c' \neq c \rightarrow isCompatible(x, dir[a][c'])$ and  $x \le state[a]$
\end{enumerate}\\
\hline
\multicolumn{1}{|c|}{Response to parent $P$ for address $a$}\\
\hline
\begin{enumerate}
\item If there is a pending response from the parent for address $a$, \ie $\neg
isWait(P, a)$, a response to the parent for $a$ can be sent only if a request
from the parent for $a$ is being handled, \ie $isHandling(P, a)$
\item If there is no pending response from the parent for address $a$, a
response to the parent for $a$ can be sent any time
\item Response \Resp{P}{a}{x} can be sent only if $x < state[a]$
% and $\forall c, x \ge dir[a][c]$
\end{enumerate}\\
\hline
\end{tabularx}
\caption{Sending a response: Preconditions}
\label{sendRespPre}
\end{figure}

\begin{figure}
\begin{tabularx}{\linewidth}{|X|}
\hline
A response for address $a$ can be sent only if the line for $a$ is present in
the current node, \ie $isPresent(a)$\\
\hline
\multicolumn{1}{|c|}{Response to child $c$ for address $a$}\\
\hline
\begin{enumerate}
\item[] On sending \Resp{c}{a}{x}, $dir[a][c]$ changes to $x$
\end{enumerate}\\
\hline
\multicolumn{1}{|c|}{Response to parent $P$ for address $a$}\\
\hline
\begin{enumerate}
\item[] On sending \Resp{P}{a}{x}, $state[a]$ changes to $x$
\end{enumerate}\\
\hline
\end{tabularx}
\caption{Sending a response: Side effects}
\label{sendRespEff}
\end{figure}

\begin{figure}
\begin{tabularx}{\linewidth}{|X|}
\hline
\multicolumn{1}{|c|}{On receiving a response from child $c$ for address $a$}\\
\hline
\begin{enumerate}
\item $dir[a][c]$ changes to $x$
\item Response \Resp{c}{a}{x} is removed
\item $isWait(c, a)$ is set \False
\end{enumerate}\\
\hline
\multicolumn{1}{|c|}{On receiving a response from parent $c$ for address $a$}\\
\hline
\begin{enumerate}
\item $state[a]$ changes to $x$
\item Response \Resp{P}{a}{x} is removed
\item $isWait(P, a)$ is set \False
\end{enumerate}\\
\hline
\end{tabularx}
\caption{Receiving a response: Effects}
\label{recvResp}
\end{figure}

@@

\begin{figure}
\begin{tabularx}{\linewidth}{|X|}
\hline
\multicolumn{1}{|c|}{Request to child $c$ for address $a$}\\
\hline
\begin{enumerate}
\item A request can be sent to child $c$ for address $a$ only if there is no
pending response from child $c$ for $a$, \ie $isWait(c, a)$
\item Request \Req{c}{a}{x} can be sent only if $x < dir[a][c]$
\end{enumerate}\\
\hline
\multicolumn{1}{|c|}{Request to parent $P$ for address $a$}\\
\hline
\begin{enumerate}
\item A request can be sent to the parent for address $a$ only if
  \begin{enumerate}
  \item there is no pending response from the parent for address $a$, \ie
  $isWait(P, a)$, and
  \item line in address $a$ is not being evicted, \ie $isEvicting(a)$
  \end{enumerate}
\item Request \Req{P}{a}{x} can be sent only if $x > state[a]$
\end{enumerate}\\
\hline
\end{tabularx}
\caption{Sending a request: Preconditions}
\label{sendReq}
\end{figure}

\begin{figure}
\begin{tabularx}{\linewidth}{|X|}
\hline
\multicolumn{1}{|X|}{\centering Request \Req{c}{a}{x} from child $c$ is being handled, \ie
$isHandling(c, a)$}\\
\hline
\begin{enumerate}
\item[] If $x \le state[a] \wedge \forall c'\neq c, isCompatible(x, dir[a][c])$,
 then a response \Resp{c}{a}{x} is sent to $c$ if $x > dir[a][c]$
%\item If $x = \Inv$ then data for address $a$ is transferred to $c$
\end{enumerate}\\
\hline
\multicolumn{1}{|c|}{Address $a$ is marked as being evicted, \ie $isEviting(a)$}\\
\hline
\begin{enumerate}
\item[] If $\forall c, dir[a][c] \le \Inv$, a response \Resp{P}{a}{\Inv} is sent
to the parent if $\Inv < state[a]$
%\item If $state[a] \in \Dirty$ then data for address $a$ is transferred to $c$
\end{enumerate}\\
\hline
\multicolumn{1}{|X|}{\centering Request \Req{P}{a}{x} from parent $P$ is being
handled, \ie $isHandling(P, a)$}\\
\hline
\begin{enumerate}
\item[] If $\forall c, dir[a][c] \le x$, a response \Resp{P}{a}{x} is sent to the
parent if $x < state[a]$
%\item If $state[a] \in \Dirty$ then data for address $a$ is transferred to the parent
\end{enumerate}\\
\hline
\end{tabularx}
\caption{Forced responses: Effects}
\label{respForReq}
\end{figure}

\begin{figure}
\begin{tabularx}{\linewidth}{|X|}
\hline
\multicolumn{1}{|c|}{Request \Req{c}{a}{x} from child $c$ is received}\\
\hline
\begin{enumerate}
\item If $a$ is present in the node, handling of request \Req{c}{a}{x} can be
started, \ie $isHandling(c, a)$ can be set \True only if
  \begin{enumerate}
  \item no request \Req{n}{a}{x} from any node $n$ for address $a$ is being
  handled, \ie $\forall n, \neg isHandling(n, a)$
  \item address $a$ is not currently being evicted, \ie $\neg isEvicting(a)$
  \end{enumerate}
\item If $a$ is not present in the node, handling of request \Req{c}{a}{x} can be
started \ie $isHandling(c, a$ can be set \True only if there is a line with
address $a'$ such that
  \begin{enumerate}
  \item no request \Req{n}{a'}{x} from any node $n$ for address $a'$ is being
  handled, \ie $\neg isHandling(n, a')$
  \item address $a'$ is not currently being evicted, \ie $\neg isEvicting(a')$
  \end{enumerate}
\end{enumerate}\\
\hline
\multicolumn{1}{|c|}{Request \Req{P}{a}{x} from parent $P$ is received}\\
\hline
\begin{enumerate}
\item[] Handling of request \Req{P}{a}{x} can be started \ie $isHandling(P, a)$
can be set \True, only if there are no pending responses from any child for
address $a$, \ie $\forall c \in Children, \neg isWait(c, a)$
\end{enumerate}\\
\hline
\end{tabularx}
\caption{Start handling a request: Preconditions}
\label{respForReqPre}
\end{figure}

\begin{figure}
\begin{tabularx}{\linewidth}{|X|}
\hline
\multicolumn{1}{|c|}{Request \Req{c}{a}{x} from child $c$ is received}\\
\hline
\begin{enumerate}
\item Handling of the request \Req{c}{a}{x} has started, \ie $isHandling(c, a)$
is set \True
\item If $a$ is not present in the node, get a line with address $a'$ such that
  \begin{enumerate}
  \item no request \Req{n}{a'}{x} from any node $n$ for address $a'$ is being
  handled, \ie $\forall n, \neg isHandling(n, a')$
  \item address $a'$ is not currently being evicted, \ie $\neg isEvicting(a')$
  \end{enumerate}
$a'$ is marked as being evicted, \ie $isEvicting(a')$ is set \True
\end{enumerate}\\
\hline
\multicolumn{1}{|c|}{Request \Req{P}{a}{x} from parent $P$ is received}\\
\hline
\begin{enumerate}
\item[] Handling of the request \Req{P}{a}{x} is started, \ie $isHandling(P, a)$ is set \True
\end{enumerate}\\
\hline
\end{tabularx}
\caption{Start handling a request: Effects}
\label{respForReqEff}
\end{figure}

\begin{figure}
\begin{tabularx}{\linewidth}{|X|}
\hline
\multicolumn{1}{|c|}{Request \Req{c}{a}{x} from child $c$ is being handled}\\
\hline
\begin{enumerate}
\item If $x \le dir[a][c]$, the request is removed
\item $isHandling(c, a)$ is set \False
\end{enumerate}\\
\hline
\multicolumn{1}{|c|}{Request \Req{P}{a}{x} from parent $P$ is being
handled}\\
\hline
\begin{enumerate}
\item If $x \ge state[a]$, the request is removed
\item $isHandling(P, a)$ is set \False
\end{enumerate}\\
\hline
\end{tabularx}
\caption{Finish handling a request: Effects}
\label{finishHandle}
\end{figure}

\begin{figure}
\begin{tabularx}{\linewidth}{|X|}
\hline
Request \Req{c}{a}{x} from child $c$ is being handled, and address $a'$ is
replaced by $a$\\
\hline
\begin{enumerate}
\item If $state[a'] <= \Inv$, address $a'$ for the line is replaced by $a$,
$state[a']$ as \Inv and $\forall c, dir(c, a)$ as \Inv
\item $a'$ is no longer being evicted, \ie set $isEvicted(a')$ to \False
\end{enumerate}\\
\hline
\end{tabularx}
\caption{Finish replacing line: Effects}
\end{figure}

\begin{figure}
\begin{tabularx}{\linewidth}{|X|}
\hline
\begin{enumerate}
\item If a request that is being handled requires $state[a]$ to be $x$ while
currently $state[a] < x$ and there is no pending response from the parent for
$a$, \ie $\neg isWait(P, a)$, a request \Req{P}{a}{x} is sent to the parent and
$isWait(P, a)$ is set \True
\item If a request that is being handled requires $dir[a][c]$ for some $c$ to
be $x$ while currently $dir[a][c] > x$ and there is no pending response from
child $c$, \ie $\neg isWait(c, a)$, a request \Req{c}{a}{x} is sent to $c$ and
$isWait(c, a)$ is set
\True
\end{enumerate}\\
\hline
\end{tabularx}
\caption{Handling of an incoming request: Effects}
\label{handling}
\end{figure}

\begin{figure}
\begin{tabularx}{\linewidth}{|X|}
\hline
\begin{enumerate}
\item The effects during sending a response or sending a request is atomic,
\ie if there is no space available in the sending buffer, then the action is
aborted without causing any effect and retried later
\item The effect of abort can be achieved by checking if the required resources
are available before starting the action 
\end{enumerate}\\
\hline
\end{tabularx}
\caption{Atomicity requirements}
\label{atomicity}
\end{figure}
